---
title: "AD Fine-Mapping and Heritability Enrichment Analysis"
subtitle: "Public Reproducible Notebook"
output: html_document
---

This notebook reproduces all main and supplementary figures from the AD fine-mapping
analysis using only public data files from the Zenodo release.

Set `DATA_DIR` below to the directory where the Zenodo files were downloaded.
All 32 files should be in a single flat directory (no subdirectories).

# Setup

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(readr)
library(patchwork)
library(ggsci)
library(ggrepel)
library(scales)
library(ComplexUpset)
library(GenomicRanges)
library(IRanges)
library(PRROC)

options(bitmapType = "cairo")                                                                                                                                                                           


DATA_DIR <- "/gpfs/commons/groups/knowles_lab/ADSP_reguloML/adfinemapping_public"
```

## Theme and Colors

```{r theme_setup}
global_text_size <- 7

theme_nature <- function(base_size = global_text_size,
                        base_family = "Helvetica") {
  theme_minimal(base_size = base_size, base_family = base_family) %+replace%
    theme(
      text = element_text(size = base_size, family = base_family, face = "plain"),
      plot.title = element_text(size = base_size, face = "plain"),
      plot.subtitle = element_text(size = base_size, face = "plain"),
      axis.title = element_text(size = base_size, face = "plain"),
      axis.text = element_text(size = base_size, face = "plain"),
      legend.title = element_text(size = base_size, face = "plain"),
      legend.text = element_text(size = base_size, face = "plain"),
      strip.text = element_text(size = base_size, face = "plain")
    )
}

npg_colors <- pal_npg()(7)

colors_vec <- c(
  'Baseline' = npg_colors[1],
  "Roadmap" = npg_colors[2],
  "Brain Omics" = npg_colors[3],
  "DeepSea" = npg_colors[4],
  'Enformer' = npg_colors[5],
  'Enformer TF/Brain Omics' = npg_colors[6],
  'SpliceAI/AlphaMissense' = npg_colors[7]
)

cohort_colors_vec <- c(
  'SUSIE' = npg_colors[1],
  "Baseline" = npg_colors[2],
  "Baseline/Omics" = npg_colors[3],
  "Baseline/Omics/DL" = npg_colors[4]
)

text_size <- 8
```

## Helper Functions

```{r helper_functions}
rename_strings_dataframe_exact <- function(input_df, column_name, mapping_df) {
  column_name <- enquo(column_name)
  input_df %>%
    rowwise() %>%
    mutate(
      replacement_index = if_else(
        any(mapping_df$Original == !!column_name),
        which(mapping_df$Original == !!column_name)[1],
        NA_integer_
      )
    ) %>%
    ungroup() %>%
    mutate(!!column_name := if_else(!is.na(replacement_index), mapping_df$Replacement[replacement_index], !!column_name)) %>%
    select(-replacement_index)
}

annotate_peak_overlap <- function(variants_df, peak_df, col_name) {
  variant_gr <- GRanges(
    seqnames = paste0("chr", variants_df$CHR),
    ranges = IRanges(start = variants_df$BP, width = 1)
  )
  peak_gr <- GRanges(
    seqnames = peak_df$chrom,
    ranges = IRanges(start = peak_df$start, end = peak_df$end)
  )
  overlaps <- overlapsAny(variant_gr, peak_gr)
  variants_df[[col_name]] <- as.integer(overlaps)
  return(variants_df)
}

get_top_enhancer_variant_matches <- function(variants_df, enhancer_df) {
  result_df <- variants_df
  result_df$enhancer_name <- NA
  result_df$TargetGene <- NA
  result_df$ABC.Score <- NA
  for(i in seq_len(nrow(variants_df))) {
    var <- variants_df[i,]
    matches <- enhancer_df$chr == paste0("chr", var$CHR) &
               enhancer_df$start <= var$BP &
               enhancer_df$end >= var$BP
    if(any(matches)) {
      matching_enhancers <- enhancer_df[matches, ]
      top_match <- matching_enhancers[which.max(matching_enhancers$ABC.Score), ]
      result_df$enhancer_name[i] <- top_match$name
      result_df$TargetGene[i] <- top_match$TargetGene
      result_df$ABC.Score[i] <- top_match$ABC.Score
    }
  }
  return(result_df)
}

get_top_placseq_variant_matches <- function(variants_df, enhancer_df, interaction_type_val) {
  result_df <- variants_df
  enhancer_df <- enhancer_df %>% dplyr::filter(interaction_type == interaction_type_val)
  result_df$TargetGene <- NA
  result_df$num_iteractions <- NA
  for(i in seq_len(nrow(variants_df))) {
    var <- variants_df[i,]
    matches <- enhancer_df$chr == paste0("chr", var$CHR) &
               enhancer_df$start <= var$BP &
               enhancer_df$end >= var$BP
    if(any(matches)) {
      matching_enhancers <- enhancer_df[matches, ]
      top_match <- matching_enhancers %>%
        dplyr::select(gene_name) %>%
        group_by(gene_name) %>%
        summarise(num_overlaps = n()) %>% slice_max(num_overlaps, n = 1) %>%
        ungroup()
      result_df$TargetGene[i] <- top_match[1,]$gene_name
      result_df$num_iteractions[i] <- top_match[1,]$num_overlaps
    }
  }
  return(result_df)
}

make_auprc_plot_90_modified <- function(data, text_position = "bottom", title = NULL, text_size = 3) {
  models <- unique(data$model)
  results <- list()
  baseline <- data %>%
    group_by(PIP_filter) %>%
    summarize(count = n()) %>%
    mutate(prop = count / sum(count)) %>%
    filter(PIP_filter == 'PIP>0.9') %>%
    pull(prop)
  for(m in models) {
    subset_data <- data %>% filter(model == m)
    pr <- PRROC::pr.curve(
      scores.class0 = subset_data %>% filter(PIP_filter == 'PIP>0.9') %>% pull(score),
      scores.class1 = subset_data %>% filter(PIP_filter == 'PIP<0.01') %>% pull(score),
      curve = TRUE
    )
    optimal_idx <- which.max((2 * pr$curve[,1] * pr$curve[,2]) / (pr$curve[,1] + pr$curve[,2]))
    results[[m]] <- data.frame(
      recall = pr$curve[,1],
      precision = pr$curve[,2],
      model = m,
      auprc = pr$auc.integral,
      optimal_cutpoint = pr$curve[optimal_idx, 3]
    )
  }
  plot_data <- bind_rows(results)
  y_min <- min(plot_data$precision, baseline) * 0.9
  y_max <- max(plot_data$precision) * 1.05
  if(text_position == "top") {
    y_range <- seq(0.9, 0.8, length.out = length(models))
    x_pos <- 0.55
  } else {
    bottom_portion <- y_min + (y_max - y_min) * 0.25
    y_range <- seq(bottom_portion, y_min + 0.05, length.out = length(models))
    x_pos <- 0.1
  }
  text_data <- plot_data %>%
    group_by(model) %>%
    summarize(auprc = dplyr::first(auprc)) %>%
    mutate(y_pos = y_range, baseline = baseline)
  p <- ggplot(plot_data) +
    geom_hline(yintercept = baseline, linetype = "dashed", color = "gray50") +
    geom_line(aes(x = recall, y = precision, color = model)) +
    theme_minimal() +
    geom_text(data = text_data,
              aes(x = x_pos, y = y_pos,
                  label = sprintf("%s AUPRC: %.3f", model, auprc),
                  color = model),
              size = text_size, hjust = 0, fontface = "bold") +
    annotate("text", x = 0.9, y = baseline + 0.02,
             label = sprintf("Baseline: %.3f", baseline),
             size = text_size - 0.5, hjust = 1, color = "gray50") +
    ggsci::scale_color_npg() +
    labs(color = "Model") +
    theme(legend.position = "bottom") +
    xlab('Recall') + ylab('Precision') +
    theme(plot.margin = unit(c(0.5, 1.5, 0.5, 0.5), "lines"))
  if (!is.null(title)) p <- p + labs(title = title)
  return(p)
}
```

## Column Mappings

```{r column_mappings}
cell_names_list_exact <- list(
  c('microglia', 'Microglia'),
  c('neuron', 'Neuron'),
  c('astrocyte', 'Astrocyte'),
  c('oligodendrocyte', 'Oligodendrocyte')
)
cell_names_exact_df <- as.data.frame(do.call(rbind, cell_names_list_exact))
colnames(cell_names_exact_df) <- c("Original", "Replacement")

gwas_names <- c('Kunkle (2019)', 'Jansen (2019)', 'Wightman (2021)', 'Bellenguez (2022)')
```

## Constants

```{r constants}
h2_kunkle_ridge <- 0.1229
M_kunkle_ridge <- 9418280
h2_bellenguez_ADSP_ridge <- 0.0618
M_bellenguez_ADSP_ridge <- 16013174
PIP_thresholds <- c(0.5, 0.6, 0.7, 0.8, 0.9, 0.95, 0.99)

count_unique_credible_sets <- function(df, pip_thresholds) {
  result <- lapply(pip_thresholds, function(threshold) {
    filtered_df <- df[df$PIP >= threshold & df$P_top_pip < 0.05, ]
    counts <- aggregate(CREDIBLE_SET ~ fine_mapping_type,
                        data = filtered_df,
                        FUN = function(x) length(unique(x)))
    counts$PIP_threshold <- threshold
    counts
  })
  do.call(rbind, result)
}

count_variants_nonCS <- function(df, pip_thresholds) {
  result <- lapply(pip_thresholds, function(threshold) {
    filtered_df <- df[df$PIP >= threshold & df$P < 0.05, ]
    counts <- filtered_df %>% group_by(fine_mapping_type) %>% tally()
    counts$PIP_threshold <- threshold
    counts
  })
  do.call(rbind, result)
}
```


# Load Data

## Per-GWAS Annotation Files

Each file contains Baseline + Glass Epigenomics + Enformer TF/Glass + ChromBPNet results.

```{r load_annotations}
annotations_kunkle <- read_delim(
  paste0(DATA_DIR, "/annotations_kunkle.tsv"),
  delim = "\t", escape_double = FALSE, trim_ws = TRUE
)

annotations_wightman <- read_delim(
  paste0(DATA_DIR, "/annotations_wightman.tsv"),
  delim = "\t", escape_double = FALSE, trim_ws = TRUE
)

annotations_bellenguez <- read_delim(
  paste0(DATA_DIR, "/annotations_bellenguez.tsv"),
  delim = "\t", escape_double = FALSE, trim_ws = TRUE
)

# Combined for supplementary figures
cell_CRE_multgwas_ldsc <- bind_rows(annotations_kunkle, annotations_wightman, annotations_bellenguez) %>%
  mutate(gwas_name = factor(gwas_name, levels = c('Kunkle (2019)', 'Wightman (2021)', 'Bellenguez (2022)')))
```

## Multivariate LDSC

```{r load_multivariate}
multivariate_with_se <- read_delim(
  paste0(DATA_DIR, "/multivariate_annotations_final.tsv"),
  delim = "\t", escape_double = FALSE, trim_ws = TRUE
)
```

## Fine-Mapping Data

```{r load_finemapping}
pip_threshold <- 0.50

zzz1 <- read_delim(paste0(DATA_DIR, "/bellenguez_susie.txt.gz"),
  "\t", escape_double = FALSE, trim_ws = TRUE) %>%
  filter((PIP >= pip_threshold) | (!str_ends(CREDIBLE_SET, ':0'))) %>%
  mutate(fine_mapping_type = 'SUSIE')

zzz2 <- read_delim(paste0(DATA_DIR, "/bellenguez_baseline.txt.gz"),
  "\t", escape_double = FALSE, trim_ws = TRUE) %>%
  filter((PIP >= pip_threshold) | (!str_ends(CREDIBLE_SET, ':0'))) %>%
  mutate(fine_mapping_type = 'Baseline')

zzz3 <- read_delim(paste0(DATA_DIR, "/bellenguez_omics.txt.gz"),
  "\t", escape_double = FALSE, trim_ws = TRUE) %>%
  filter((PIP >= pip_threshold) | (!str_ends(CREDIBLE_SET, ':0'))) %>%
  mutate(fine_mapping_type = 'Baseline/Omics')

zzz4 <- read_delim(paste0(DATA_DIR, "/bellenguez_omics_dl.txt.gz"),
  "\t", escape_double = FALSE, trim_ws = TRUE) %>%
  filter((PIP >= pip_threshold) | (!str_ends(CREDIBLE_SET, ':0'))) %>%
  mutate(fine_mapping_type = 'Baseline/Omics/DL')
```

## Process Fine-Mapping Counts

```{r process_finemapping}
susie_finemap_singleton <- zzz1 %>% filter(str_ends(CREDIBLE_SET, ':0'))
susie_finemap_CS <- zzz1 %>% filter(!str_ends(CREDIBLE_SET, ':0'))
baseline_finemap_singleton <- zzz2 %>% filter(str_ends(CREDIBLE_SET, ':0'))
baseline_finemap_CS <- zzz2 %>% filter(!str_ends(CREDIBLE_SET, ':0'))
omics_finemap_singleton <- zzz3 %>% filter(str_ends(CREDIBLE_SET, ':0'))
omics_finemap_CS <- zzz3 %>% filter(!str_ends(CREDIBLE_SET, ':0'))
omics_dl_finemap_singleton <- zzz4 %>% filter(str_ends(CREDIBLE_SET, ':0'))
omics_dl_finemap_CS <- zzz4 %>% filter(!str_ends(CREDIBLE_SET, ':0'))

finemapped_combined_CS <- bind_rows(susie_finemap_CS, baseline_finemap_CS, omics_finemap_CS, omics_dl_finemap_CS)
finemapped_combined_singleton <- bind_rows(susie_finemap_singleton, baseline_finemap_singleton, omics_finemap_singleton, omics_dl_finemap_singleton)

CS_statistics <- finemapped_combined_CS %>%
  group_by(fine_mapping_type, CREDIBLE_SET) %>%
  summarise(PIP = max(PIP), CS_size = n(), P_top_pip = P[which.max(PIP)], .groups = "drop")

CS_statistics_summary <- count_unique_credible_sets(CS_statistics, PIP_thresholds) %>%
  mutate(PIP_threshold = as.factor(PIP_threshold)) %>%
  mutate(variant_type = 'Credible Sets') %>%
  mutate(n = CREDIBLE_SET)

singleton_statistics_summary <- count_variants_nonCS(finemapped_combined_singleton, PIP_thresholds) %>%
  mutate(PIP_threshold = as.factor(PIP_threshold)) %>%
  mutate(variant_type = 'Non Credible Set Variants')

finemapped_combined_thresholds_combined <- bind_rows(CS_statistics_summary, singleton_statistics_summary)
finemapped_combined_thresholds_combined$fine_mapping_type <- factor(
  finemapped_combined_thresholds_combined$fine_mapping_type,
  levels = c('SUSIE', 'Baseline', 'Baseline/Omics', 'Baseline/Omics/DL')
)
```

## PRS Data (Pre-computed)

```{r load_prs}
summary_SNPS <- read_delim(paste0(DATA_DIR, "/prs_auc_summary.tsv"),
  delim = "\t", escape_double = FALSE, trim_ws = TRUE)

PRS_results <- read_delim(paste0(DATA_DIR, "/prs_odds_ratios.tsv"),
  delim = "\t", escape_double = FALSE, trim_ws = TRUE)

prevalence_by_quintile <- read_delim(paste0(DATA_DIR, "/prs_prevalence_by_decile.tsv"),
  delim = "\t", escape_double = FALSE, trim_ws = TRUE)
```

## Variant Annotations

```{r load_variant_annotations}
top_omics_dl_finemapped_variants_annotated <- read_tsv(
  paste0(DATA_DIR, "/top_omics_dl_finemapped_variants_annotated.tsv.gz"),
  show_col_types = FALSE
)

parse_favor_gene <- function(info) {
  stringr::str_extract(info, "^[A-Za-z0-9._-]+")
}

top_omics_dl_finemapped_variants_annotated <- top_omics_dl_finemapped_variants_annotated %>%
  dplyr::mutate(
    consensus_category = dplyr::case_when(
      favor_category == "exonic" ~ "exonic",
      vep_category == "ncrna_exon" ~ "ncrna_exon",
      vep_category %in% c("UTR5", "UTR3") & favor_category %in% c("upstream", "intergenic", "intronic") ~ vep_category,
      vep_category == "splicing" & favor_category == "intronic" ~ "splicing",
      favor_category == "intergenic" & vep_category == "intronic" ~ "intronic",
      favor_category == "intergenic" & vep_category == "upstream" ~ "upstream",
      TRUE ~ favor_category
    ),
    consensus_gene = dplyr::case_when(
      consensus_category == "intergenic" ~ "",
      consensus_category %in% c("ncrna_exon", "UTR5", "UTR3", "upstream", "splicing") &
        !is.na(vep_gene_symbol) & vep_gene_symbol != "" ~ vep_gene_symbol,
      !is.na(favor_raw_info) & favor_raw_info != "" ~ parse_favor_gene(favor_raw_info),
      TRUE ~ dplyr::coalesce(vep_gene_symbol, "")
    ),
    meta_category = dplyr::case_when(
      `GenecodeComprehensiveExonicCategory.String` == "nonsynonymous SNV" ~ "Coding - Missense",
      `GenecodeComprehensiveExonicCategory.String` == "synonymous SNV" ~ "Coding - Synonymous",
      `GenecodeComprehensiveExonicCategory.String` == "stopgain" ~ "Stop Gained",
      consensus_category == "exonic" ~ "Coding - Other",
      consensus_category == "intronic" ~ "Intronic",
      consensus_category == "intergenic" ~ "Intergenic",
      TRUE ~ "Miscellaneous"
    )
  )

exonic_variants <- top_omics_dl_finemapped_variants_annotated %>%
  dplyr::filter(FunseqDescription.String == 'coding')

noncoding_variants <- top_omics_dl_finemapped_variants_annotated %>%
  dplyr::filter(FunseqDescription.String == 'noncoding')
```

## ATAC Peaks, ABC, PLAC-seq

```{r load_peaks_abc}
microglia_ATAC <- read.table(paste0(DATA_DIR, '/PU1_promoter_enhancer_plus_atac.bed'), sep = '\t')
colnames(microglia_ATAC) <- c('chrom', 'start', 'end')

oligodendrocyte_ATAC <- read.table(paste0(DATA_DIR, '/Olig2_promoter_enhancer_plus_atac.bed'), sep = '\t')
colnames(oligodendrocyte_ATAC) <- c('chrom', 'start', 'end')

astrocyte_ATAC <- read.table(paste0(DATA_DIR, '/LHX2_promoter_enhancer_plus_atac.bed'), sep = '\t')
colnames(astrocyte_ATAC) <- c('chrom', 'start', 'end')

neuron_ATAC <- read.table(paste0(DATA_DIR, '/NeuN_promoter_enhancer_plus_atac.bed'), sep = '\t')
colnames(neuron_ATAC) <- c('chrom', 'start', 'end')

noncoding_variants_peak <- annotate_peak_overlap(noncoding_variants, microglia_ATAC, 'peak_microglia_ATAC')
noncoding_variants_peak <- annotate_peak_overlap(noncoding_variants_peak, oligodendrocyte_ATAC, 'peak_oligodendrocyte_ATAC')
noncoding_variants_peak <- annotate_peak_overlap(noncoding_variants_peak, astrocyte_ATAC, 'peak_astrocyte_ATAC')
noncoding_variants_peak <- annotate_peak_overlap(noncoding_variants_peak, neuron_ATAC, 'peak_neuron_ATAC')

ABC_microglia <- read_delim(paste0(DATA_DIR, "/microglia_EnhancerPredictionsAllPutative.tsv.gz"),
  delim = "\t", escape_double = FALSE, trim_ws = TRUE)
ABC_oligodendrocyte <- read_delim(paste0(DATA_DIR, "/oligodendrocyte_EnhancerPredictionsAllPutative.tsv.gz"),
  delim = "\t", escape_double = FALSE, trim_ws = TRUE)
ABC_astrocyte <- read_delim(paste0(DATA_DIR, "/astrocyte_EnhancerPredictionsAllPutative.tsv.gz"),
  delim = "\t", escape_double = FALSE, trim_ws = TRUE)
ABC_neuron <- read_delim(paste0(DATA_DIR, "/neuron_EnhancerPredictionsAllPutative.tsv.gz"),
  delim = "\t", escape_double = FALSE, trim_ws = TRUE)

noncoding_variants_peak <- get_top_enhancer_variant_matches(noncoding_variants_peak, ABC_microglia) %>%
  dplyr::rename(microglia_enhancer_name = enhancer_name, microglia_TargetGene = TargetGene, microglia_ABC.Score = ABC.Score)
noncoding_variants_peak <- get_top_enhancer_variant_matches(noncoding_variants_peak, ABC_oligodendrocyte) %>%
  dplyr::rename(oligodendrocyte_enhancer_name = enhancer_name, oligodendrocyte_TargetGene = TargetGene, oligodendrocyte_ABC.Score = ABC.Score)
noncoding_variants_peak <- get_top_enhancer_variant_matches(noncoding_variants_peak, ABC_astrocyte) %>%
  dplyr::rename(astrocyte_enhancer_name = enhancer_name, astrocyte_TargetGene = TargetGene, astrocyte_ABC.Score = ABC.Score)
noncoding_variants_peak <- get_top_enhancer_variant_matches(noncoding_variants_peak, ABC_neuron) %>%
  dplyr::rename(neuron_enhancer_name = enhancer_name, neuron_TargetGene = TargetGene, neuron_ABC.Score = ABC.Score)

placseq_interactions <- read.table(paste0(DATA_DIR, '/placseq_interactions_hg38.tsv'), sep = '\t', header = TRUE)

noncoding_variants_peak <- get_top_placseq_variant_matches(noncoding_variants_peak, placseq_interactions, 'PU1_enhancer_interactions') %>%
  dplyr::rename(microglia_placseq_TargetGene = TargetGene, microglia_placseq_interactions = num_iteractions)
noncoding_variants_peak <- get_top_placseq_variant_matches(noncoding_variants_peak, placseq_interactions, 'NeuN_enhancer_interactions') %>%
  dplyr::rename(neuron_placseq_TargetGene = TargetGene, neuron_placseq_interactions = num_iteractions)
noncoding_variants_peak <- get_top_placseq_variant_matches(noncoding_variants_peak, placseq_interactions, 'Olig2_enhancer_interactions') %>%
  dplyr::rename(oligodendrocyte_placseq_TargetGene = TargetGene, oligodendrocyte_placseq_interactions = num_iteractions)

noncoding_variants_peak <- noncoding_variants_peak %>%
  mutate(microglia_evidence = peak_microglia_ATAC) %>%
  mutate(oligodendrocyte_evidence = peak_oligodendrocyte_ATAC) %>%
  mutate(neuron_evidence = peak_neuron_ATAC) %>%
  mutate(astrocyte_evidence = peak_astrocyte_ATAC)
```

## Build Upset Data

```{r build_upset}
upset_variant_list <- bind_rows(
  exonic_variants %>% dplyr::select(VariantVcf.String, CHR, BP, A1, A2, meta_category,
    GenecodeComprehensiveCategory.String, GenecodeComprehensiveExonicCategory.String),
  noncoding_variants_peak %>% dplyr::select(VariantVcf.String, CHR, BP, A1, A2, meta_category,
    GenecodeComprehensiveCategory.String, GenecodeComprehensiveExonicCategory.String,
    microglia_evidence, oligodendrocyte_evidence, neuron_evidence, astrocyte_evidence)
)

upset_variant_list <- upset_variant_list %>%
  dplyr::mutate(
    `Microglia CRE`       = dplyr::coalesce(microglia_evidence == 1, FALSE),
    `Astrocyte CRE`       = dplyr::coalesce(astrocyte_evidence == 1, FALSE),
    `Neuron CRE`          = dplyr::coalesce(neuron_evidence == 1, FALSE),
    `Oligodendrocyte CRE` = dplyr::coalesce(oligodendrocyte_evidence == 1, FALSE),
    `Intronic`            = meta_category == "Intronic",
    `Miscellaneous`       = meta_category == "Miscellaneous",
    `Coding - Synonymous` = meta_category == "Coding - Synonymous",
    `Coding - Missense`   = meta_category == "Coding - Missense",
    `Stop Gained`         = meta_category == "Stop Gained",
    `Intergenic`          = meta_category == "Intergenic"
  ) %>%
  dplyr::select(VariantVcf.String, `Microglia CRE`, `Astrocyte CRE`, `Neuron CRE`, `Oligodendrocyte CRE`,
    `Intronic`, `Miscellaneous`, `Coding - Synonymous`, `Coding - Missense`, `Stop Gained`, `Intergenic`)

region_names <- colnames(upset_variant_list)[-1]
```

## caQTL Data

```{r load_caqtl}
TF_microglia <- c('ETS2', 'FLI1', 'SPI1', 'IRF8', 'PRDM1', 'CEBPA', 'CEBPB', 'CEBPD', 'CEBPE',
    'KLF11', 'KLF2', 'TFEC', 'BACH1', 'BATF', 'BATF3', 'RUNX1', 'RUNX2', 'RUNX3',
    'LYL1', 'TAL1', 'ELK3', 'MEF2C', 'CREB3L2', 'ATF4', 'MAFB', 'SALL1', 'eGFP-SALL1',
    'SMAD5', 'MEF2A', 'MEF2B', 'SMAD2', 'USF1', 'STAT3', 'eGFP-MAFG', 'NFYB', 'NRF1', 'CREB1', 'IRF1',
    'SOX9', 'SP1', 'IRF4', '3xFLAG-NFYC', 'ZNF384', 'MITF', 'JUNB', 'MAFF')

caQTL_chrombpnet_peak_centered <- read_delim(paste0(DATA_DIR, "/chrombpnet_microglia_peak_centered_PIP_090.tsv"),
  delim = "\t", escape_double = FALSE, trim_ws = TRUE) %>%
  dplyr::select(SNP, PIP_filter, ChromBPNet_microglia_score = log_counts_diff_chrombpnet_abs)

caQTL_enformer_peak_centered <- read_delim(paste0(DATA_DIR, "/enformer_microglia_peak_centered_PIP_090.tsv"),
  delim = "\t", escape_double = FALSE, trim_ws = TRUE) %>%
  dplyr::select(SNP, PIP_filter, Enformer_TF_score = log_counts_diff_enformer_TF_abs,
    Enformer_DNASE_monocyte_score = log_counts_diff_enformer_monocyte_DNASE_abs)

motifBreakR <- read_delim(paste0(DATA_DIR, "/mb_output.tsv"),
  delim = "\t", escape_double = FALSE, trim_ws = TRUE) %>%
  mutate(motifBreakR_score = abs(alleleDiff)) %>%
  filter(geneSymbol %in% TF_microglia) %>%
  group_by(SNP_id) %>%
  summarize(motifBreakR_score = mean(motifBreakR_score)) %>%
  ungroup() %>%
  dplyr::select(SNP = SNP_id, motifBreakR_score)

rsid_list <- caQTL_chrombpnet_peak_centered %>% dplyr::select(SNP)
motifBreakR <- motifBreakR %>% left_join(rsid_list, by = 'SNP')

caQTL_enformer_predictions_peaks <- caQTL_chrombpnet_peak_centered %>%
  inner_join(caQTL_enformer_peak_centered, by = c('SNP', 'PIP_filter')) %>%
  inner_join(motifBreakR, by = c('SNP')) %>%
  pivot_longer(cols = ChromBPNet_microglia_score:motifBreakR_score, names_to = 'Annotation', values_to = 'score') %>%
  mutate(model = case_when(
    Annotation == 'Enformer_TF_score' ~ 'Enformer TF',
    Annotation == 'Enformer_DNASE_monocyte_score' ~ 'Enformer DNASE',
    Annotation == 'ChromBPNet_microglia_score' ~ 'ChromBPNet ATAC',
    Annotation == 'motifBreakR_score' ~ 'motifBreakR'
  ))

caQTL_chrombpnet_no_distance <- read_delim(paste0(DATA_DIR, "/chrombpnet_microglia_no_distance_PIP_090.tsv"),
  delim = "\t", escape_double = FALSE, trim_ws = TRUE) %>%
  dplyr::select(SNP, PIP_filter, ChromBPNet_microglia_score = log_counts_diff_chrombpnet_abs)

caQTL_enformer_no_distance <- read_delim(paste0(DATA_DIR, "/enformer_microglia_no_distance_PIP_090.tsv"),
  delim = "\t", escape_double = FALSE, trim_ws = TRUE) %>%
  dplyr::select(SNP, PIP_filter, Enformer_TF_score = log_counts_diff_enformer_TF_abs,
    Enformer_DNASE_monocyte_score = log_counts_diff_enformer_monocyte_DNASE_abs)

caQTL_enformer_predictions_nonpeaks <- caQTL_chrombpnet_no_distance %>%
  inner_join(caQTL_enformer_no_distance, by = c('SNP', 'PIP_filter')) %>%
  pivot_longer(cols = ChromBPNet_microglia_score:Enformer_DNASE_monocyte_score, names_to = 'Annotation', values_to = 'score') %>%
  mutate(model = case_when(
    Annotation == 'Enformer_TF_score' ~ 'Enformer TF',
    Annotation == 'Enformer_DNASE_monocyte_score' ~ 'Enformer DNASE',
    Annotation == 'ChromBPNet_microglia_score' ~ 'ChromBPNet ATAC'
  ))
```


# Figure 2: Cell Type Enrichment + caQTL AUPRC

## caQTL AUPRC Plots

```{r figure2_caqtl}
plot_caqtl_density_peak <- ggplot(caQTL_enformer_predictions_peaks, aes(x = score, fill = PIP_filter)) +
  geom_density(alpha = 0.3) +
  facet_wrap(~model, scales = 'free') +
  scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
                labels = trans_format("log10", math_format(10^.x))) +
  scale_fill_npg() + theme_nature() +
  xlab('VEP Score') + ylab('Density') + labs(fill = "PIP Threshold") +
  theme(legend.position = "bottom") + ggtitle('Peak Centered Scores')

plot_caqtl_density_nonpeak <- ggplot(caQTL_enformer_predictions_nonpeaks, aes(x = score, fill = PIP_filter)) +
  geom_density(alpha = 0.3) +
  facet_wrap(~model, scales = 'free', ncol = 2) +
  scale_x_log10(labels = trans_format("log10", math_format(10^.x))) +
  scale_fill_npg() + theme_nature() +
  xlab('VEP Score') + ylab('Density') + labs(fill = "PIP Threshold") +
  theme(legend.position = "bottom") + ggtitle('Variant Centered Scores')

plot_auprc_cqtl_peak <- make_auprc_plot_90_modified(caQTL_enformer_predictions_peaks, text_position = "bottom", title = "Peak Centered Scores", text_size = 2) + theme_nature()
plot_auprc_cqtl_nonpeak <- make_auprc_plot_90_modified(caQTL_enformer_predictions_nonpeaks, text_position = "top", title = "Variant Centered Scores", text_size = 2) + theme_nature()

plot_caqtl_density_combined <- (plot_caqtl_density_peak | plot_caqtl_density_nonpeak) +
  plot_layout(guides = "collect") & theme(legend.position = "bottom")

plot_auprc_cqtl_combined <- (plot_auprc_cqtl_peak | plot_auprc_cqtl_nonpeak) +
  plot_layout(guides = "collect") & theme(legend.position = "bottom")

plot_composite_density_auprc_microglia_qtl <- plot_caqtl_density_combined / plot_auprc_cqtl_combined
plot_composite_density_auprc_microglia_qtl <- plot_composite_density_auprc_microglia_qtl &
  theme(legend.key.size = unit(0.8, "lines"),
        legend.spacing.x = unit(0.2, 'cm'),
        legend.spacing.y = unit(0.1, 'cm'),
        legend.margin = margin(t = 0, r = 0, b = 0, l = 0))

plot_caqtl_density_peak
plot_caqtl_density_nonpeak
plot_auprc_cqtl_peak
plot_auprc_cqtl_nonpeak
```

## Cell Type Enrichment

```{r figure2_cell_type}
cell_filter <- c('microglia', 'neuron', 'astrocyte', 'oligodendrocyte')
genomic_region_filter <- c('enhancer_promoter_union_atac_500')
cell_order <- c('Microglia', 'Neuron', 'Oligodendrocyte', 'Astrocyte')
cell_order_reverse <- c('Astrocyte', 'Oligodendrocyte', 'Neuron', 'Microglia')

cell_type_df <- bind_rows(
  annotations_kunkle %>%
    filter(data_source == 'Glass Epigenomics') %>%
    filter(cell_name %in% cell_filter) %>%
    filter(assay %in% genomic_region_filter) %>%
    mutate(gwas = gwas_names[1]),
  annotations_wightman %>%
    filter(data_source == 'Glass Epigenomics') %>%
    filter(cell_name %in% cell_filter) %>%
    filter(assay %in% genomic_region_filter) %>%
    mutate(gwas = gwas_names[3]),
  annotations_bellenguez %>%
    filter(data_source == 'Glass Epigenomics') %>%
    filter(cell_name %in% cell_filter) %>%
    filter(assay %in% genomic_region_filter) %>%
    mutate(gwas = gwas_names[4])
) %>%
  rename_strings_dataframe_exact(cell_name, cell_names_exact_df) %>%
  mutate(cell_name = factor(cell_name, levels = cell_order)) %>%
  mutate(h2_CI_low = Prop._h2 - 1.96 * Prop._h2_std_error) %>%
  mutate(h2_CI_high = Prop._h2 + 1.96 * Prop._h2_std_error) %>%
  mutate(gwas = factor(gwas, levels = gwas_names))

plot_prop_SNPs <- ggplot(cell_type_df, aes(x = gwas, y = Prop._SNPs, fill = cell_name)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  coord_flip() + xlab("GWAS") + ylab('Proportion of SNPs in Cell Type CRE') +
  labs(fill = "Cell Type") + scale_fill_npg(breaks = cell_order_reverse) +
  theme_nature() + theme(legend.position = "none")

plot_prop_h2 <- ggplot(cell_type_df, aes(x = gwas, y = Prop._h2, fill = cell_name)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  coord_flip() + xlab(NULL) + ylab('Proportion of Heritability') +
  labs(fill = "Cell Type") +
  geom_errorbar(aes(ymin = h2_CI_low, ymax = h2_CI_high), width = .2, position = position_dodge(.9)) +
  scale_fill_npg(breaks = cell_order_reverse) + theme_nature() +
  theme(axis.text.y = element_blank(), axis.title.y = element_blank(), legend.position = "none")

plot_enrichment <- ggplot(cell_type_df, aes(x = gwas, y = Enrichment, fill = cell_name)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  coord_flip() + xlab(NULL) + ylab('Enrichment of Heritability') +
  labs(fill = "Cell Type") +
  geom_errorbar(aes(ymin = CI_low, ymax = CI_high), width = .2, position = position_dodge(.9)) +
  scale_fill_npg(breaks = cell_order_reverse) + theme_nature() +
  theme(axis.text.y = element_blank(), axis.title.y = element_blank())

cell_type_plot <- (plot_prop_SNPs | plot_prop_h2 | plot_enrichment) +
  plot_layout(guides = "collect") & theme(legend.position = "bottom")

plot_prop_SNPs
plot_prop_h2
plot_enrichment
```

## Figure 2 Composite

```{r figure2_composite, fig.width=7, fig.height=8.3}
figure2 <- (cell_type_plot / plot_composite_density_auprc_microglia_qtl) + plot_layout(height = c(1, 2))

figure2

```


# Figure 3: Pareto Plot + Multivariate tau*

## Pareto Plot

```{r figure3_pareto}
cols_filter <- c(
  'chrombpnet_microglia-enhancer_promoter_union_atac_500-microglia_combined-zscore_1_1',
  'chrombpnet_neuron-enhancer_promoter_union_atac_500-neuron_combined-zscore_1_1',
  'astrocyte-enhancer_promoter_union_atac_500-astrocyte-zscore_1_1',
  'astrocyte-enhancer_promoter_union_atac_500-astrocyte-zscore_2_1',
  'astrocyte-enhancer_promoter_union_atac_500-astrocyte-zscore_3_1',
  'microglia-enhancer_promoter_union_atac_500-microglia-zscore_1_1',
  'microglia-enhancer_promoter_union_atac_500-microglia-zscore_2_1',
  'microglia-enhancer_promoter_union_atac_500-microglia-zscore_3_1',
  'neuron-enhancer_promoter_union_atac_500-neuron-zscore_1_1',
  'neuron-enhancer_promoter_union_atac_500-neuron-zscore_2_1',
  'neuron-enhancer_promoter_union_atac_500-neuron-zscore_3_1',
  'oligodendrocyte-enhancer_promoter_union_atac_500-oligodendrocyte-zscore_1_1',
  'oligodendrocyte-enhancer_promoter_union_atac_500-oligodendrocyte-zscore_2_1',
  'oligodendrocyte-enhancer_promoter_union_atac_500-oligodendrocyte-zscore_3_1',
  'astrocyte_enhancer_promoter_union_atac_500_1',
  'microglia_enhancer_promoter_union_atac_500_1',
  'neuron_enhancer_promoter_union_atac_500_1',
  'oligodendrocyte_enhancer_promoter_union_atac_500_1'
)

cell_type_df_subset <- annotations_bellenguez %>%
  filter(Category %in% cols_filter) %>%
  mutate(annotation_type = case_when(
    (data_source == 'Enformer Tensorflow TF/Glass Epigenomics' & pct_category == 'zscore_1') ~ 'TF ZS>1',
    (data_source == 'Enformer Tensorflow TF/Glass Epigenomics' & pct_category == 'zscore_2') ~ 'TF ZS>2',
    (data_source == 'Enformer Tensorflow TF/Glass Epigenomics' & pct_category == 'zscore_3') ~ 'TF ZS>3',
    (data_source == 'ChromBPNet Enformer Tensorflow TF/Glass Epigenomics' & pct_category == 'zscore_1') ~ 'TF+C ZS>1',
    (data_source == 'Glass Epigenomics') ~ 'CRE'
  )) %>%
  mutate(cell_name_formal = case_when(
    cell_name == 'astrocyte' ~ 'Astrocyte',
    cell_name == 'microglia' ~ 'Microglia',
    cell_name == 'neuron' ~ 'Neuron',
    cell_name == 'oligodendrocyte' ~ 'Oligodendrocyte',
    cell_name == 'chrombpnet_microglia' ~ 'Microglia',
    cell_name == 'chrombpnet_neuron' ~ 'Neuron'
  )) %>%
  mutate(cell_name_formal = factor(cell_name_formal, levels = cell_order)) %>%
  filter(cell_name_formal != 'Astrocyte')

cell_type_df_subset_nochrombpnet <- cell_type_df_subset %>% filter(annotation_type != 'TF+C ZS>1')
cell_type_df_subset_chrombpnet <- cell_type_df_subset %>% filter(annotation_type == 'TF+C ZS>1')
cell_type_df_subset_nochrombpnet$annotation_type <- factor(cell_type_df_subset_nochrombpnet$annotation_type,
  levels = c("CRE", "TF ZS>1", "TF ZS>2", "TF ZS>3"))
cell_type_df_subset_nochrombpnet <- cell_type_df_subset_nochrombpnet %>% arrange(cell_name, annotation_type)

pareto_text_size <- 3

pareto_plot <- ggplot(cell_type_df_subset_nochrombpnet,
  aes(x = Enrichment, y = Prop._h2, color = cell_name_formal, group = cell_name_formal)) +
  geom_path() +
  geom_point(size = 3, alpha = 0.8) +
  geom_text_repel(aes(label = annotation_type), box.padding = 0.8, point.padding = 0.5,
    segment.color = 'grey50', show.legend = FALSE, size = pareto_text_size,
    max.overlaps = 30, force = 10, force_pull = 0.2, direction = "both") +
  geom_point(data = cell_type_df_subset_chrombpnet,
    aes(x = Enrichment, y = Prop._h2, color = cell_name_formal), size = 3) +
  geom_text_repel(data = cell_type_df_subset_chrombpnet,
    aes(x = Enrichment, y = Prop._h2, label = annotation_type, color = cell_name_formal),
    box.padding = 1.0, point.padding = 0.8, segment.color = 'grey50', show.legend = FALSE,
    size = pareto_text_size, max.overlaps = 30, nudge_y = 0.05, force = 15, direction = "both") +
  theme_nature() + scale_fill_npg() + scale_color_npg() +
  labs(title = "Enrichment and Proportion of Heritability of AD GWAS within Brain Cell Types",
    x = "Enrichment of Heritability", y = "Proportion of Heritability", color = "Cell Type") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "darkgrey") +
  geom_vline(xintercept = 0, linetype = "dashed", color = "darkgrey") +
  theme(legend.position = "bottom")
```

## Multivariate tau*

```{r figure3_multivariate}
annotation_order <- c('Baseline', 'Roadmap', 'Enformer', 'Brain Omics/Enformer/ChromBPNet', 'SpliceAI/AlphaMissense')

annotation_colors_vec <- c(
  'Baseline' = npg_colors[1],
  "Roadmap" = npg_colors[2],
  "Enformer" = npg_colors[3],
  "Brain Omics/Enformer/ChromBPNet" = npg_colors[4],
  "SpliceAI/AlphaMissense" = npg_colors[5]
)

multivariate_filtered <- multivariate_with_se %>%
  filter(Enrichment_p < 0.05 & tau_star > 0 & Enrichment > 5.0) %>%
  mutate(name = case_when(
    id == 'TSS_Hoffman.extend.500' ~ 'TSS_Hoffman',
    id == 'Coding_UCSC.extend.500' ~ 'Coding_UCSC',
    id == 'alleleage' ~ 'Allele Age',
    id == 'Ancient_Sequence_Age_Human_Promoter.extend.500' ~ 'Ancient Sequence Age Promoter',
    id == 'BLUEPRINT_FE_META_TISSUE_DNAMETH_MaxCPP' ~ 'DNA Methylation Multi-Tissue',
    id == 'BLUEPRINT_FE_META_TISSUE_H3K27ac_MaxCPP' ~ 'H3K27ac Multi-Tissue',
    id == 'Conserved_Mammal_phastCons46way.extend.500' ~ 'Conserved PhastCons46way',
    id == 'microglia_enhancer_intersect_atac' ~ 'Microglia-ATAC/Enhancer',
    id == 'diff_32_ENCFF748RWG_zscore_1' ~ 'CHIP:POLR2AphosS5-Spleen (ZS1)',
    id == 'diff_32_ENCFF472KMT_zscore' ~ 'CHIP:ATF2 (ZS3)',
    id == 'diff_32_ENCFF476YVH_zscore_2' ~ 'CHIP:H3K27me3:Liver (ZS2)',
    id == 'Human_Enhancer_Villar_Species_Enhancer_Count' ~ 'Enhancer Species-Villar',
    id == 'H3K27ac_PGC2' ~ 'H3K27ac PGC2',
    id == 'H3K4me3_Trynka' ~ 'H3K4me3 Trynka',
    id == 'Conserved_Primate_phastCons46way' ~ 'Conserved Primate phastCons',
    id == 'DGF_ENCODE.extend.500' ~ 'DGF ENCODE',
    id == 'Human_Enhancer_Villar.extend.500' ~ 'Human Enhancer Villar',
    id == 'E051-H3K27ac.imputed.narrowPeak' ~ 'CD34+ H3K27ac',
    id == 'diff_32_ENCFF316ZVD_zscore_1' ~ 'CHIP:H3K4me3:PBMC (ZS1)',
    id == 'diff_32_ENCFF572IZL_zscore_2' ~ 'CHIP:H4K20me1:B cell (ZS2)',
    id == 'diff_32_ENCFF991YWM_zscore' ~ 'CHIP:DNASE:RPMI-7951 (ZS3)',
    id == 'diff_32_ENCFF890NAY_zscore_2' ~ 'CHIP:H3K27ac:DOHH2 (ZS2)',
    id == 'chrombpnet_microglia-enhancer_promoter_union_atac_500-microglia_combined-zscore_1' ~ 'Microglia/Enformer CRE (ZS1)',
    id == 'chrombpnet_neuron-enhancer_promoter_union_atac_500-neuron_combined-zscore_3' ~ 'Neuron/Enformer CRE (ZS3)',
    id == 'TSS_Hoffman' ~ 'TSS Hoffman',
    id == 'all_celltype_spliceai_05_or_alphamissense' ~ 'SpliceAI/AlphaMissense',
    id == 'GTEx_FE_META_TISSUE_GE_MaxCPP' ~ 'GTEx_FE_META_TISSUE_GE_MaxCPP',
    id == 'Vahedi_Tcell_SE' ~ 'Vahedi Tcell SE',
    id == 'Enhancer_Hoffman.extend.500' ~ 'Enhancer Hoffman',
    id == 'Promoter_UCSC' ~ 'Promoter UCSC',
    id == 'BivFlnk' ~ 'BivFlnk',
    id == 'E030-H3K4me1.imputed.narrowPeak' ~ 'Neutrophils H3K4me1',
    id == 'E031-DNase.imputed.narrowPeak' ~ 'B Cell DNase',
    id == 'E051-DNase.imputed.narrowPeak' ~ 'Hematopoietic Stem Cell DNase',
    id == 'BivFlnk.extend.500' ~ 'BivFlnk Extended '
  )) %>%
  mutate(annotation = case_when(
    id %in% c('TSS_Hoffman.extend.500', 'Coding_UCSC.extend.500', 'alleleage',
              'Ancient_Sequence_Age_Human_Promoter.extend.500', 'BLUEPRINT_FE_META_TISSUE_DNAMETH_MaxCPP',
              'BLUEPRINT_FE_META_TISSUE_H3K27ac_MaxCPP', 'Conserved_Mammal_phastCons46way.extend.500',
              'Human_Enhancer_Villar_Species_Enhancer_Count', 'H3K27ac_PGC2', 'H3K4me3_Trynka',
              'Conserved_Primate_phastCons46way', 'DGF_ENCODE.extend.500', 'Human_Enhancer_Villar.extend.500',
              'TSS_Hoffman', 'GTEx_FE_META_TISSUE_GE_MaxCPP', 'Vahedi_Tcell_SE',
              'Enhancer_Hoffman.extend.500', 'Promoter_UCSC', 'BivFlnk', 'BivFlnk.extend.500') ~ 'Baseline',
    id %in% c('E051-H3K27ac.imputed.narrowPeak', 'E030-H3K4me1.imputed.narrowPeak',
              'E031-DNase.imputed.narrowPeak', 'E051-DNase.imputed.narrowPeak') ~ 'Roadmap',
    id %in% c('diff_32_ENCFF748RWG_zscore_1', 'diff_32_ENCFF472KMT_zscore', 'diff_32_ENCFF476YVH_zscore_2',
              'diff_32_ENCFF316ZVD_zscore_1', 'diff_32_ENCFF572IZL_zscore_2', 'diff_32_ENCFF991YWM_zscore',
              'diff_32_ENCFF890NAY_zscore_2') ~ 'Enformer',
    id == 'microglia_enhancer_intersect_atac' ~ 'Brain Omics',
    id %in% c('chrombpnet_microglia-enhancer_promoter_union_atac_500-microglia_combined-zscore_1',
              'chrombpnet_neuron-enhancer_promoter_union_atac_500-neuron_combined-zscore_3') ~ 'Brain Omics/Enformer/ChromBPNet',
    id == 'all_celltype_spliceai_05_or_alphamissense' ~ 'SpliceAI/AlphaMissense'
  )) %>%
  mutate(annotation = factor(annotation, levels = annotation_order))

bellenguez_plot_enrichment_h2 <- ggplot(multivariate_filtered, aes(x = reorder(name, Enrichment), y = Enrichment, fill = annotation)) +
  geom_bar(stat = "identity") + coord_flip() +
  scale_fill_manual(values = annotation_colors_vec) +
  xlab("Annotation") + ylab('Enrichment of Heritability') + labs(fill = "Annotation Category") +
  geom_errorbar(aes(ymin = CI_low, ymax = CI_high), width = .2, position = position_dodge(.9)) +
  theme_nature() + theme(legend.position = "bottom")

bellenguez_plot_prop_h2 <- ggplot(multivariate_filtered, aes(x = reorder(name, Prop._h2), y = Prop._h2, fill = annotation)) +
  geom_bar(stat = "identity") + coord_flip() +
  scale_fill_manual(values = annotation_colors_vec) +
  xlab("Annotation") + ylab('Proportion of Heritability') + labs(fill = "Annotation Category") +
  geom_errorbar(aes(ymin = CI_prop_h2_low, ymax = CI_prop_h2_high), width = .2, position = position_dodge(.9)) +
  theme_nature() + theme(legend.position = "none")

bellenguez_plot_prop_SNPs <- ggplot(multivariate_filtered, aes(x = reorder(name, Prop._SNPs), y = Prop._SNPs, fill = annotation)) +
  geom_bar(stat = "identity") + coord_flip() +
  scale_fill_manual(values = annotation_colors_vec) +
  xlab("Annotation") + ylab('Proportion of Variants in Annotation') + labs(fill = "Annotation Category") +
  theme_nature() + theme(legend.position = "none")

multivariate_filtered_plot <- multivariate_filtered %>%
  filter(tau_star_pvalue < 0.05) %>%
  mutate(tau_star_low = tau_star - 1.96 * tau_star_SE) %>%
  mutate(tau_star_high = tau_star + 1.96 * tau_star_SE)

bellenguez_plot_tau_star <- ggplot(multivariate_filtered_plot, aes(x = reorder(name, tau_star), y = tau_star, fill = annotation)) +
  geom_bar(stat = "identity") + coord_flip() +
  scale_fill_manual(values = annotation_colors_vec) +
  xlab("Annotation") + ylab('tau* Coefficient') + labs(fill = "Annotation Category") +
  geom_errorbar(aes(ymin = tau_star_low, ymax = tau_star_high), width = .2, position = position_dodge(.9)) +
  theme_nature() + theme(legend.position = "none")

composite_plot_enrichment <- (bellenguez_plot_prop_SNPs | bellenguez_plot_prop_h2) / (bellenguez_plot_enrichment_h2 | bellenguez_plot_tau_star)

pareto_plot
bellenguez_plot_prop_SNPs
bellenguez_plot_prop_h2
bellenguez_plot_enrichment_h2
bellenguez_plot_tau_star
```

## Figure 3 Composite

```{r figure3_composite, fig.width=7, fig.height=8.3}
figure3 <- (pareto_plot / composite_plot_enrichment) + plot_layout(height = c(1, 2.5))
figure3
```


# Figure 4: Fine-Mapping + PRS

## Fine-Map Counts

```{r figure4_finemap}
finemap_bar_CS <- ggplot(data = finemapped_combined_thresholds_combined %>% filter(variant_type == 'Credible Sets'),
  aes(x = PIP_threshold, y = n, fill = fine_mapping_type)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  xlab('Number of CS with\nMax PIP Above Threshold') +
  scale_fill_manual(values = cohort_colors_vec) +
  ylab('Number of Credible Sets') +
  guides(fill = guide_legend(title = "Annotations")) +
  theme_nature()

finemap_bar_singleton <- ggplot(data = finemapped_combined_thresholds_combined %>% filter(variant_type == 'Non Credible Set Variants'),
  aes(x = PIP_threshold, y = n, fill = fine_mapping_type)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  xlab('Number of Singleton Variants\nwith PIP Above Threshold') +
  scale_fill_manual(values = cohort_colors_vec) +
  ylab('Number of Variants') +
  guides(fill = guide_legend(title = "Annotations")) +
  theme_nature()

credible_set_count_plot <- (finemap_bar_CS | finemap_bar_singleton) +
  plot_layout(guides = "collect") & ggplot2::theme(legend.position = 'bottom')

finemap_bar_CS
finemap_bar_singleton
```

## Upset Plot

```{r figure4_upset}
upset_finemapped <- ComplexUpset::upset(upset_variant_list,
  region_names, min_size = 2, name = 'Genomic Region',
  queries = list(
    upset_query(set = 'Microglia CRE', color = npg_colors[1], fill = npg_colors[1]),
    upset_query(set = 'Neuron CRE', color = npg_colors[2], fill = npg_colors[2]),
    upset_query(set = 'Oligodendrocyte CRE', color = npg_colors[3], fill = npg_colors[3]),
    upset_query(set = 'Astrocyte CRE', color = npg_colors[4], fill = npg_colors[4]),
    upset_query(set = 'Intronic', color = npg_colors[5], fill = npg_colors[5]),
    upset_query(intersect = c('Microglia CRE'), fill = npg_colors[1],
      only_components = c('intersections_matrix', 'Intersection size'))
  )
) + theme_nature()
```

## PRS AUC

```{r figure4_prs}
summary_SNPS_figure <- summary_SNPS %>%
  mutate(annotation = case_when(
    cohort == 'bellenguez_susie' ~ 'SUSIE',
    cohort == 'bellenguez_baseline' ~ 'Baseline',
    cohort == 'bellenguez_omics' ~ 'Baseline/Omics',
    cohort == 'bellenguez_omics_dl' ~ 'Baseline/Omics/DL'
  )) %>%
  mutate(auc_low = auc - 1.96 * auc_se) %>%
  mutate(auc_high = auc + 1.96 * auc_se) %>%
  mutate(threshold = as.factor(threshold)) %>%
  group_by(predicted_ancestry, annotation) %>% slice_max(auc, n = 1) %>% ungroup() %>%
  mutate(ancestry_label = case_when(
    predicted_ancestry == 'AFR' ~ 'African American',
    predicted_ancestry == 'EUR' ~ 'European',
    predicted_ancestry == 'AMR' ~ 'Hispanic',
    predicted_ancestry == 'SAS' ~ 'South Asian',
    predicted_ancestry == 'EAS' ~ 'East Asian'
  )) %>%
  mutate(annotation = factor(annotation, levels = c("SUSIE", "Baseline", "Baseline/Omics", "Baseline/Omics/DL"))) %>%
  mutate(ancestry_label = factor(ancestry_label, levels = c("African American", "South Asian", "East Asian", "Hispanic", "European")))

PRS_AUC_Plot_modified <- ggplot(summary_SNPS_figure, aes(x = auc, y = ancestry_label, color = annotation)) +
  geom_point(position = position_dodge(width = 0.5)) +
  geom_errorbarh(aes(xmin = auc_low, xmax = auc_high), position = position_dodge(width = 0.5), height = 0.2) +
  scale_color_manual(values = cohort_colors_vec) +
  theme_nature() + labs(x = "PRS AUROC", y = "Ancestry", color = "Annotation") +
  theme(legend.position = "bottom") + xlim(0.45, 0.65)
```

## Odds Ratio Plot

```{r figure4_odds_ratio}
PRS_results_modified <- PRS_results %>%
  mutate(ancestry_label = case_when(
    predicted_ancestry == 'AFR' ~ 'African American',
    predicted_ancestry == 'EUR' ~ 'European',
    predicted_ancestry == 'AMR' ~ 'Hispanic',
    predicted_ancestry == 'SAS' ~ 'South Asian',
    predicted_ancestry == 'EAS' ~ 'East Asian'
  )) %>%
  mutate(annotation_label = case_when(
    annotation == 'BL' ~ 'Baseline',
    annotation == 'BL/Omics' ~ 'Baseline/Omics',
    annotation == 'BL/Omics/DL' ~ 'Baseline/Omics/DL',
    annotation == 'SUSIE' ~ 'SUSIE'
  )) %>%
  mutate(annotation_label = factor(annotation_label, levels = c("SUSIE", "Baseline", "Baseline/Omics", "Baseline/Omics/DL"))) %>%
  mutate(ancestry_label = factor(ancestry_label, levels = c("African American", "South Asian", "East Asian", "Hispanic", "European")))

odds_ratio_plot <- ggplot(PRS_results_modified, aes(x = estimate, y = ancestry_label, color = annotation_label)) +
  geom_point(position = position_dodge(width = 0.5)) +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), position = position_dodge(width = 0.5), height = 0.2) +
  theme_nature() + scale_color_manual(values = cohort_colors_vec) +
  labs(x = "Odds Ratio", y = "Ancestry", color = "Annotation") +
  theme(legend.position = "bottom")
```

## Figure 4 Composite

```{r figure4_composite, fig.width=7, fig.height=8.3}
PRS_composite <- (PRS_AUC_Plot_modified | odds_ratio_plot) +
  patchwork::plot_layout(guides = 'collect') & ggplot2::theme(legend.position = 'bottom')

upset_finemapped
PRS_AUC_Plot_modified
odds_ratio_plot

figure4 <- (credible_set_count_plot) / (upset_finemapped) / PRS_composite
figure4 <- figure4 + plot_layout(height = c(1, 1.75, 1))
figure4
```


# Supplementary Figures

## Helper: Cell Type CRE Enrichment Plot

```{r supp_helper}
make_cell_type_supp_plot <- function(cell_CRE_data, cell_type_val, cell_type_val_lower) {
  pct_category_val <- c('zscore_1', 'none', 'All')

  cell_type_df <- cell_CRE_data %>%
    dplyr::filter(cell_type %in% cell_type_val) %>%
    dplyr::filter(pct_category %in% pct_category_val) %>%
    dplyr::filter(!str_detect(Category, 'ATAC_H3K27ac_H3K4me3')) %>%
    mutate(annotation = case_when(
      str_detect(Category, 'enhancer_promoter_union_atac_500-all') ~ paste0(cell_type_val, '\nCRE+Enformer\n(All TF)'),
      str_detect(Category, 'chrombpnet') ~ paste0(cell_type_val, '\nCRE+Enformer+ChromBPNet\n(', cell_type_val, ' TF)'),
      str_detect(Category, paste0('enhancer_promoter_union_atac_500-', cell_type_val_lower)) ~ paste0(cell_type_val, '\nCRE+Enformer\n(', cell_type_val, ' TF)'),
      str_detect(Category, 'enhancer_promoter_union_atac_500_1$') ~ paste0(cell_type_val, '\nCRE'),
      str_detect(Category, 'enhancer_promoter_union_atac_500-none') ~ paste0(cell_type_val, '\nCRE')
    )) %>%
    mutate(annotation = factor(annotation, levels = c(
      paste0(cell_type_val, '\nCRE'),
      paste0(cell_type_val, '\nCRE+Enformer\n(All TF)'),
      paste0(cell_type_val, '\nCRE+Enformer+ChromBPNet\n(', cell_type_val, ' TF)'),
      paste0(cell_type_val, '\nCRE+Enformer\n(', cell_type_val, ' TF)')
    )))

  plot_prop_SNPs <- ggplot(cell_type_df, aes(x = gwas_name, y = Prop._SNPs, fill = annotation)) +
    geom_bar(stat = "identity", position = position_dodge()) + coord_flip() +
    xlab("GWAS") + ylab('Proportion of SNPs in Annotation') + labs(fill = "Annotation") +
    scale_fill_npg() + theme_nature() + theme(legend.position = "none")

  plot_prop_h2 <- ggplot(cell_type_df, aes(x = gwas_name, y = Prop._h2, fill = annotation)) +
    geom_bar(stat = "identity", position = position_dodge()) +
    geom_errorbar(aes(ymin = h2_CI_low, ymax = h2_CI_high), width = .2, position = position_dodge(.9)) +
    coord_flip() + xlab(NULL) + ylab('Proportion of Heritability') + labs(fill = "Annotation") +
    scale_fill_npg() + theme_nature() +
    theme(axis.text.y = element_blank(), axis.title.y = element_blank(), legend.position = "none")

  plot_enrichment <- ggplot(cell_type_df, aes(x = gwas_name, y = Enrichment, fill = annotation)) +
    geom_bar(stat = "identity", position = position_dodge()) +
    geom_errorbar(aes(ymin = CI_low, ymax = CI_high), width = .2, position = position_dodge(.9)) +
    coord_flip() + xlab(NULL) + ylab('Enrichment of Heritability') + labs(fill = "Annotation") +
    scale_fill_npg() + theme_nature() +
    theme(axis.text.y = element_blank(), axis.title.y = element_blank())

  (plot_prop_SNPs | plot_prop_h2 | plot_enrichment) +
    plot_layout(guides = "collect") & theme(legend.position = "bottom")
}
```

## Supplementary Figure 1: Microglia

```{r supp_fig1, fig.width=7, fig.height=4}
make_cell_type_supp_plot(cell_CRE_multgwas_ldsc, 'Microglia', 'microglia')
```

## Supplementary Figure 2: Astrocyte

```{r supp_fig2, fig.width=7, fig.height=4}
make_cell_type_supp_plot(cell_CRE_multgwas_ldsc, 'Astrocyte', 'astrocyte')
```

## Supplementary Figure 3: Neuron

```{r supp_fig3, fig.width=7, fig.height=4}
make_cell_type_supp_plot(cell_CRE_multgwas_ldsc, 'Neuron', 'neuron')
```

## Supplementary Figure 4: Oligodendrocyte

```{r supp_fig4, fig.width=7, fig.height=4}
make_cell_type_supp_plot(cell_CRE_multgwas_ldsc, 'Oligodendrocyte', 'oligodendrocyte')
```

## Helper: Cell Type ZS Enrichment Plot

```{r supp_zs_helper}
make_cell_type_zs_plot <- function(cell_CRE_data, cell_type_val, cell_type_val_lower, TF_filter_val) {
  data_source_filter <- c('Glass Epigenomics', 'Enformer Tensorflow TF/Glass Epigenomics')
  pct_category_val <- c('zscore_1', 'zscore_2', 'zscore_3', 'none', 'All')

  cell_type_df <- cell_CRE_data %>%
    dplyr::filter(data_source %in% data_source_filter) %>%
    dplyr::filter(cell_type %in% cell_type_val) %>%
    dplyr::filter(TF %in% TF_filter_val) %>%
    dplyr::filter(pct_category %in% pct_category_val) %>%
    mutate(annotation = case_when(
      pct_category == 'zscore_1' ~ paste0(cell_type_val, '\nCRE+Enformer\n(ZS1)'),
      pct_category == 'zscore_2' ~ paste0(cell_type_val, '\nCRE+Enformer\n(ZS2)'),
      pct_category == 'zscore_3' ~ paste0(cell_type_val, '\nCRE+Enformer\n(ZS3)'),
      pct_category == 'none' ~ paste0(cell_type_val, '\nCRE'),
      pct_category == 'All' ~ paste0(cell_type_val, '\nCRE')
    )) %>%
    mutate(annotation = factor(annotation, levels = c(
      paste0(cell_type_val, '\nCRE'),
      paste0(cell_type_val, '\nCRE+Enformer\n(ZS1)'),
      paste0(cell_type_val, '\nCRE+Enformer\n(ZS2)'),
      paste0(cell_type_val, '\nCRE+Enformer\n(ZS3)')
    )))

  plot_prop_SNPs <- ggplot(cell_type_df, aes(x = gwas_name, y = Prop._SNPs, fill = annotation)) +
    geom_bar(stat = "identity", position = position_dodge()) + coord_flip() +
    xlab("GWAS") + ylab('Proportion of SNPs in Annotation') + labs(fill = "Annotation") +
    scale_fill_npg() + theme_nature() + theme(legend.position = "none")

  plot_prop_h2 <- ggplot(cell_type_df, aes(x = gwas_name, y = Prop._h2, fill = annotation)) +
    geom_bar(stat = "identity", position = position_dodge()) +
    geom_errorbar(aes(ymin = h2_CI_low, ymax = h2_CI_high), width = .2, position = position_dodge(.9)) +
    coord_flip() + xlab(NULL) + ylab('Proportion of Heritability') + labs(fill = "Annotation") +
    scale_fill_npg() + theme_nature() +
    theme(axis.text.y = element_blank(), axis.title.y = element_blank(), legend.position = "none")

  plot_enrichment <- ggplot(cell_type_df, aes(x = gwas_name, y = Enrichment, fill = annotation)) +
    geom_bar(stat = "identity", position = position_dodge()) +
    geom_errorbar(aes(ymin = CI_low, ymax = CI_high), width = .2, position = position_dodge(.9)) +
    coord_flip() + xlab(NULL) + ylab('Enrichment of Heritability') + labs(fill = "Annotation") +
    scale_fill_npg() + theme_nature() +
    theme(axis.text.y = element_blank(), axis.title.y = element_blank())

  (plot_prop_SNPs | plot_prop_h2 | plot_enrichment) +
    plot_layout(guides = "collect") & theme(legend.position = "bottom")
}
```

## Supplementary Figure 5: Microglia ZS Enrichment

```{r supp_fig5, fig.width=7, fig.height=4}
make_cell_type_zs_plot(cell_CRE_multgwas_ldsc, 'Microglia', 'microglia', c('microglia', 'none'))
```

## Supplementary Figure 6: Astrocyte ZS Enrichment

```{r supp_fig6, fig.width=7, fig.height=4}
make_cell_type_zs_plot(cell_CRE_multgwas_ldsc, 'Astrocyte', 'astrocyte', c('astrocyte', 'none'))
```

## Supplementary Figure 7: Neuron ZS Enrichment

```{r supp_fig7, fig.width=7, fig.height=4}
make_cell_type_zs_plot(cell_CRE_multgwas_ldsc, 'Neuron', 'neuron', c('neuron', 'none'))
```

## Supplementary Figure 8: Oligodendrocyte ZS Enrichment

```{r supp_fig8, fig.width=7, fig.height=4}
make_cell_type_zs_plot(cell_CRE_multgwas_ldsc, 'Oligodendrocyte', 'oligodendrocyte', c('oligodendrocyte', 'none'))
```

## Credible Set Summary Statistics (Auxiliary)

```{r cs_summary_stats}
omics_finemap_singleton_filtered <- finemapped_combined_singleton %>%
  filter(fine_mapping_type == 'Baseline/Omics') %>% filter(P < 0.05)
omics_finemap_CS_top_filtered <- finemapped_combined_CS %>%
  filter(fine_mapping_type == 'Baseline/Omics') %>%
  group_by(CREDIBLE_SET) %>% slice_max(PIP, n = 1) %>% ungroup() %>% filter(P < 0.05)

omics_dl_finemap_singleton_filtered <- finemapped_combined_singleton %>%
  filter(fine_mapping_type == 'Baseline/Omics/DL') %>% filter(P < 0.05)
omics_dl_finemap_CS_top_filtered <- finemapped_combined_CS %>%
  filter(fine_mapping_type == 'Baseline/Omics/DL') %>%
  group_by(CREDIBLE_SET) %>% slice_max(PIP, n = 1) %>% ungroup() %>% filter(P < 0.05)

susie_finemap_singleton_filtered <- finemapped_combined_singleton %>%
  filter(fine_mapping_type == 'SUSIE') %>% filter(P < 0.05)
susie_finemap_CS_top_filtered <- finemapped_combined_CS %>%
  filter(fine_mapping_type == 'SUSIE') %>%
  group_by(CREDIBLE_SET) %>% slice_max(PIP, n = 1) %>% ungroup() %>% filter(P < 0.05)

baseline_finemap_singleton_filtered <- finemapped_combined_singleton %>%
  filter(fine_mapping_type == 'Baseline') %>% filter(P < 0.05)
baseline_finemap_CS_top_filtered <- finemapped_combined_CS %>%
  filter(fine_mapping_type == 'Baseline') %>%
  group_by(CREDIBLE_SET) %>% slice_max(PIP, n = 1) %>% ungroup() %>% filter(P < 0.05)

finemapped_combined_CS_filtered <- bind_rows(susie_finemap_CS_top_filtered,
                                              baseline_finemap_CS_top_filtered,
                                              omics_finemap_CS_top_filtered,
                                              omics_dl_finemap_CS_top_filtered)

finemapped_combined_singleton_filtered <- bind_rows(susie_finemap_singleton_filtered,
                                                     baseline_finemap_singleton_filtered,
                                                     omics_finemap_singleton_filtered,
                                                     omics_dl_finemap_singleton_filtered)

num_CS_all <- finemapped_combined_CS_filtered %>%
  group_by(fine_mapping_type) %>% summarise(num_cs = n()) %>%
  mutate(annotation = factor(fine_mapping_type, levels = c("SUSIE", "Baseline", "Baseline/Omics", "Baseline/Omics/DL")))

num_singletons_all <- finemapped_combined_singleton_filtered %>%
  group_by(fine_mapping_type) %>% summarise(num_cs = n()) %>%
  mutate(annotation = factor(fine_mapping_type, levels = c("SUSIE", "Baseline", "Baseline/Omics", "Baseline/Omics/DL")))

num_CS_size_plot <- ggplot(num_CS_all, aes(x = annotation, y = num_cs, fill = annotation)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  xlab('Finemapping Method') + scale_fill_manual(values = cohort_colors_vec) +
  ylab('Total Number of CS') + guides(fill = guide_legend(title = "Annotations")) +
  theme_nature()

num_singleton_size_plot <- ggplot(num_singletons_all, aes(x = annotation, y = num_cs, fill = annotation)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  xlab('Finemapping Method') + scale_fill_manual(values = cohort_colors_vec) +
  ylab('Total Number of Singleton Variants (PIP > 0.50)') + guides(fill = guide_legend(title = "Annotations")) +
  theme_nature()

num_CS_size_plot
num_singleton_size_plot

count_cs_singleton_all <- (num_CS_size_plot) / (num_singleton_size_plot) +
  patchwork::plot_layout(guides = 'collect') & ggplot2::theme(legend.position = 'bottom')

count_cs_singleton_all
```

## Mean Number of Variants in CS (Auxiliary)

```{r cs_mean_size}
count_mean_variant_size_credible_sets <- function(df, pip_thresholds) {
  result <- lapply(pip_thresholds, function(threshold) {
    filtered_df <- df[df$max_pip >= threshold & df$P_top_pip < 0.05, ]
    counts <- filtered_df %>% group_by(fine_mapping_type) %>% summarise(mean_num_variants = mean(num_variants))
    counts$PIP_threshold <- threshold
    counts
  })
  do.call(rbind, result)
}

PIP_thresholds_supp <- c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 0.95, 0.99)

finemapped_combined_CS_top_PIP_P <- finemapped_combined_CS %>%
  group_by(fine_mapping_type, CREDIBLE_SET) %>%
  slice_max(PIP, n = 1, with_ties = FALSE) %>% ungroup() %>%
  dplyr::select(fine_mapping_type, CREDIBLE_SET, P_top_pip = P)

finemapped_CS_statistics_pvalues <- finemapped_combined_CS_top_PIP_P %>%
  dplyr::select(CREDIBLE_SET, fine_mapping_type, P_top_pip) %>% unique()

finemapped_CS_statistics <- finemapped_combined_CS %>%
  group_by(CREDIBLE_SET, fine_mapping_type) %>%
  summarise(num_variants = n(), max_pip = max(PIP)) %>% ungroup() %>%
  left_join(finemapped_CS_statistics_pvalues, by = c('CREDIBLE_SET', 'fine_mapping_type'))

mean_CS_size <- count_mean_variant_size_credible_sets(finemapped_CS_statistics, PIP_thresholds_supp) %>%
  mutate(PIP_threshold = as.factor(PIP_threshold))

mean_CS_size$fine_mapping_type <- factor(mean_CS_size$fine_mapping_type,
  levels = c('SUSIE', 'Baseline', 'Baseline/Omics', 'Baseline/Omics/DL'))

mean_CS_size_plot <- ggplot(mean_CS_size, aes(x = PIP_threshold, y = mean_num_variants, fill = fine_mapping_type)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  xlab('All CS with\nMax PIP Above Threshold') +
  scale_fill_manual(values = cohort_colors_vec) +
  ylab('Mean Number of Variants per CS') +
  guides(fill = guide_legend(title = "Annotations")) +
  theme_nature()

mean_CS_size_plot
```

## Supplementary Figure 9: CS Size Scatterplots

```{r supp_fig9, fig.width=5, fig.height=10}
summary_finemapped_SNPs <- finemapped_combined_CS %>%
  group_by(fine_mapping_type, CREDIBLE_SET) %>%
  summarise(num_variants = n(), max_PIP = max(PIP)) %>%
  ungroup() %>%
  dplyr::filter(max_PIP > 0.10)

summary_finemapped_SNPs_loci <- summary_finemapped_SNPs %>%
  tidyr::separate(CREDIBLE_SET, into = c('chr', 'bp_range', 'idx'), sep = ':') %>%
  mutate(locus = paste0(chr, ':', bp_range)) %>%
  dplyr::select(-idx, -bp_range, -chr) %>%
  group_by(fine_mapping_type, locus) %>%
  mutate(rank_val = rank(-max_PIP, ties.method = "random"))

zz1_susie <- summary_finemapped_SNPs_loci %>%
  dplyr::filter(fine_mapping_type == 'SUSIE') %>%
  dplyr::rename(num_variants_susie = num_variants)

zz1_baseline <- summary_finemapped_SNPs_loci %>%
  dplyr::filter(fine_mapping_type == 'Baseline') %>%
  dplyr::rename(num_variants_baseline = num_variants)

zz1_baseline_omics <- summary_finemapped_SNPs_loci %>%
  dplyr::filter(fine_mapping_type == 'Baseline/Omics') %>%
  dplyr::rename(num_variants_baseline_omics = num_variants)

zz1_baseline_omics_dl <- summary_finemapped_SNPs_loci %>%
  dplyr::filter(fine_mapping_type == 'Baseline/Omics/DL') %>%
  dplyr::rename(num_variants_baseline_omics_dl = num_variants)

z_susie_vs_dl <- zz1_baseline_omics_dl %>%
  inner_join(zz1_susie, by = c('locus', 'rank_val')) %>%
  mutate(dl = case_when(
    num_variants_baseline_omics_dl == num_variants_susie ~ 'TIED',
    num_variants_baseline_omics_dl > num_variants_susie ~ 'DL',
    num_variants_baseline_omics_dl < num_variants_susie ~ 'SUSIE'
  ))

z_baseline_vs_dl <- zz1_baseline_omics_dl %>%
  inner_join(zz1_baseline, by = c('locus', 'rank_val')) %>%
  mutate(dl = case_when(
    num_variants_baseline_omics_dl == num_variants_baseline ~ 'TIED',
    num_variants_baseline_omics_dl > num_variants_baseline ~ 'DL',
    num_variants_baseline_omics_dl < num_variants_baseline ~ 'Baseline'
  ))

z_baseline_omics_vs_dl <- zz1_baseline_omics_dl %>%
  inner_join(zz1_baseline_omics, by = c('locus', 'rank_val')) %>%
  mutate(dl = case_when(
    num_variants_baseline_omics_dl == num_variants_baseline_omics ~ 'TIED',
    num_variants_baseline_omics_dl > num_variants_baseline_omics ~ 'DL',
    num_variants_baseline_omics_dl < num_variants_baseline_omics ~ 'Omics'
  ))

z_susie_vs_dl %>% group_by(dl) %>% tally()
z_baseline_vs_dl %>% group_by(dl) %>% tally()
z_baseline_omics_vs_dl %>% group_by(dl) %>% tally()

p1_susie_vs_dl <- ggplot(z_susie_vs_dl, aes(y = num_variants_susie, x = num_variants_baseline_omics_dl)) +
  geom_point() + geom_abline(slope = 1, intercept = 0) + theme_nature() +
  ylab('Number of Variants in CS (SuSiE)') +
  xlab('Number of Variants in CS (Baseline+Omics+DL)') +
  xlim(0, 60) + ylim(0, 60)

p2_baseline_vs_dl <- ggplot(z_baseline_vs_dl, aes(y = num_variants_baseline, x = num_variants_baseline_omics_dl)) +
  geom_point() + geom_abline(slope = 1, intercept = 0) + theme_nature() +
  ylab('Number of Variants in CS (Baseline)') +
  xlab('Number of Variants in CS (Baseline+Omics+DL)') +
  xlim(0, 60) + ylim(0, 60)

p3_omics_vs_dl <- ggplot(z_baseline_omics_vs_dl, aes(y = num_variants_baseline_omics, x = num_variants_baseline_omics_dl)) +
  geom_point() + geom_abline(slope = 1, intercept = 0) + theme_nature() +
  ylab('Number of Variants in CS (Baseline+Omics)') +
  xlab('Number of Variants in CS (Baseline+Omics+DL)') +
  xlim(0, 60) + ylim(0, 60)

p1_susie_vs_dl
p2_baseline_vs_dl
p3_omics_vs_dl

p4_combined_all_methods <- (p1_susie_vs_dl / p2_baseline_vs_dl / p3_omics_vs_dl) +
  plot_annotation(title = 'Comparison of the Number of Variants in Matched CS') +
  theme_nature()

p4_combined_all_methods
```

## Supplementary Figure 10: Finemapped Variant Overlap

```{r supp_fig10}
finemapped_single_variants_PIP_50_long <- bind_rows(
  finemapped_combined_CS %>% filter(PIP > 0.50 & P < 0.05),
  finemapped_combined_singleton %>% filter(PIP > 0.50 & P < 0.05)
) %>%
  dplyr::select(SNP, fine_mapping_type) %>%
  mutate(membership = TRUE)

finemapped_single_variants_PIP_50_combined <- finemapped_single_variants_PIP_50_long %>%
  pivot_wider(names_from = fine_mapping_type, values_from = membership, values_fill = FALSE)

upset_annotation_names <- colnames(finemapped_single_variants_PIP_50_combined)[-1]

overlap_finemapped_variants_upsetplot_complex <- ComplexUpset::upset(
  finemapped_single_variants_PIP_50_combined,
  upset_annotation_names,
  min_size = 1,
  name = 'Annotation Type',
  queries = list(
    upset_query(set = 'SUSIE', color = npg_colors[1], fill = npg_colors[1]),
    upset_query(set = 'Baseline', color = npg_colors[2], fill = npg_colors[2]),
    upset_query(set = 'Baseline/Omics', color = npg_colors[3], fill = npg_colors[3]),
    upset_query(set = 'Baseline/Omics/DL', color = npg_colors[4], fill = npg_colors[4]),
    upset_query(intersect = c('SUSIE'), color = npg_colors[1], fill = npg_colors[1],
                only_components = c('intersections_matrix', 'Intersection size')),
    upset_query(intersect = c('Baseline'), color = npg_colors[2], fill = npg_colors[2],
                only_components = c('intersections_matrix', 'Intersection size')),
    upset_query(intersect = c('Baseline/Omics'), color = npg_colors[3], fill = npg_colors[3],
                only_components = c('intersections_matrix', 'Intersection size')),
    upset_query(intersect = c('Baseline/Omics/DL'), color = npg_colors[4], fill = npg_colors[4],
                only_components = c('intersections_matrix', 'Intersection size'))
  )
) + theme_nature()

overlap_finemapped_variants_upsetplot_complex
```

## Supplementary Figures 11-15: PRS by Ancestry

```{r supp_prs_setup}
summary_SNPS_all_plot <- summary_SNPS %>%
  mutate(annotation = case_when(
    cohort == 'bellenguez_susie' ~ 'SUSIE',
    cohort == 'bellenguez_baseline' ~ 'Baseline',
    cohort == 'bellenguez_omics' ~ 'Baseline/Omics',
    cohort == 'bellenguez_omics_dl' ~ 'Baseline/Omics/DL'
  )) %>%
  mutate(annotation = factor(annotation, levels = c("SUSIE", "Baseline", "Baseline/Omics", "Baseline/Omics/DL")))
```

## Supplementary Figure 11: PRS AUC - African American Ancestry

```{r supp_fig11, fig.width=7, fig.height=5}
PRS_african_american_plot <- ggplot(summary_SNPS_all_plot %>% filter(predicted_ancestry == 'AFR'),
  aes(x = auc, y = threshold, color = annotation)) +
  geom_point(position = position_dodge(width = 0.5)) +
  geom_errorbarh(aes(xmin = auc_low, xmax = auc_high), position = position_dodge(width = 0.5), height = 0.2) +
  scale_y_discrete(limits = rev(levels(summary_SNPS$threshold))) +
  scale_color_manual(values = cohort_colors_vec) +
  theme_nature() + labs(x = "AUC", y = "Threshold", color = "Annotation") +
  theme(legend.position = "bottom") + ggtitle('African American Ancestry') + xlim(0.40, 0.65)

PRS_african_american_plot
```

## Supplementary Figure 12: PRS AUC - Hispanic Ancestry

```{r supp_fig12, fig.width=7, fig.height=5}
PRS_hispanic_plot <- ggplot(summary_SNPS_all_plot %>% filter(predicted_ancestry == 'AMR'),
  aes(x = auc, y = threshold, color = annotation)) +
  geom_point(position = position_dodge(width = 0.5)) +
  geom_errorbarh(aes(xmin = auc_low, xmax = auc_high), position = position_dodge(width = 0.5), height = 0.2) +
  scale_y_discrete(limits = rev(levels(summary_SNPS$threshold))) +
  scale_color_manual(values = cohort_colors_vec) +
  theme_nature() + labs(x = "AUC", y = "Threshold", color = "Annotation") +
  theme(legend.position = "bottom") + ggtitle('Hispanic Ancestry') + xlim(0.40, 0.65)

PRS_hispanic_plot
```

## Supplementary Figure 13: PRS AUC - European Ancestry

```{r supp_fig13, fig.width=7, fig.height=5}
PRS_european_plot <- ggplot(summary_SNPS_all_plot %>% filter(predicted_ancestry == 'EUR'),
  aes(x = auc, y = threshold, color = annotation)) +
  geom_point(position = position_dodge(width = 0.5)) +
  geom_errorbarh(aes(xmin = auc_low, xmax = auc_high), position = position_dodge(width = 0.5), height = 0.2) +
  scale_y_discrete(limits = rev(levels(summary_SNPS$threshold))) +
  scale_color_manual(values = cohort_colors_vec) +
  theme_nature() + labs(x = "AUC", y = "Threshold", color = "Annotation") +
  theme(legend.position = "bottom") + ggtitle('European Ancestry') + xlim(0.40, 0.65)

PRS_european_plot
```

## Supplementary Figure 14: PRS AUC - East Asian Ancestry

```{r supp_fig14, fig.width=7, fig.height=5}
PRS_east_asian_plot <- ggplot(summary_SNPS_all_plot %>% filter(predicted_ancestry == 'EAS'),
  aes(x = auc, y = threshold, color = annotation)) +
  geom_point(position = position_dodge(width = 0.5)) +
  geom_errorbarh(aes(xmin = auc_low, xmax = auc_high), position = position_dodge(width = 0.5), height = 0.2) +
  scale_y_discrete(limits = rev(levels(summary_SNPS$threshold))) +
  scale_color_manual(values = cohort_colors_vec) +
  theme_nature() + labs(x = "AUC", y = "Threshold", color = "Annotation") +
  theme(legend.position = "bottom") + ggtitle('East Asian Ancestry') + xlim(0.40, 0.65)

PRS_east_asian_plot
```

## Supplementary Figure 15: PRS AUC - South Asian Ancestry

```{r supp_fig15, fig.width=7, fig.height=5}
PRS_south_asian_plot <- ggplot(summary_SNPS_all_plot %>% filter(predicted_ancestry == 'SAS'),
  aes(x = auc, y = threshold, color = annotation)) +
  geom_point(position = position_dodge(width = 0.5)) +
  geom_errorbarh(aes(xmin = auc_low, xmax = auc_high), position = position_dodge(width = 0.5), height = 0.2) +
  scale_y_discrete(limits = rev(levels(summary_SNPS$threshold))) +
  scale_color_manual(values = cohort_colors_vec) +
  theme_nature() + labs(x = "AUC", y = "Threshold", color = "Annotation") +
  theme(legend.position = "bottom") + ggtitle('South Asian Ancestry') + xlim(0.40, 0.65)

PRS_south_asian_plot
```

## PRS Prevalence by Decile (All Ancestries)

```{r prs_prevalence, fig.width=10, fig.height=6}
prevalence_by_quintile <- prevalence_by_quintile %>%
  mutate(annotation = factor(annotation, levels = c('SUSIE', 'BL', 'BL/Omics', 'BL/Omics/DL')))

ggplot(prevalence_by_quintile, aes(x = score_quintile_factor, y = prevalence, fill = annotation)) +
  geom_col(position = "dodge") +
  facet_wrap(~predicted_ancestry) +
  scale_fill_npg() + theme_minimal() +
  xlab('Decile') + ylab('Prevalence')
```
